---
title: "StatsHW2"
Authors/Collaborators: Ahmed Mostafa, Hamed Asadollahi, Mostafa Awid 
output: html_document
---
we are going to use igraph library
```{r}
library(igraph) # Load the igraph package
```

simulating node and edges creation
```{r}
generateGraph_singleOutDegree <- function(gamma=0.5){
g1 <- graph( edges=c(1,2, 2,3, 3,4, 4,1 ), n=4, directed=T ) 
maxNV = 10000
nV = 4
while(nV<maxNV)
{
  y = runif(1)
  g1 = add.vertices(g1,c(1))
  if (y <= gamma)
  {
    toV = sample(1:nV,1)
    g1 = add.edges(g1,c(nV+1,toV))
  }else
  {
    indegree = degree(g1, mode = c("in"))
    pToBeCopied = indegree/nV
    pToBeCopied = head(pToBeCopied,-1)
    toV = sample(1:nV,size=1,prob=pToBeCopied)
    g1 = add.edges(g1,c(nV+1,toV))
    #print("Y")
    #print(toV)
    
  }
  nV=nV+1
}
return(g1)
}

generateGraph_threeOutDegree <- function(gamma=0.5){
g1 <- graph.full(4,directed = T)
maxNV = 10000
nV = 4
while(nV<maxNV)
{
  y = runif(1)
  g1 = add.vertices(g1,c(1))
  if (y <= gamma)
  {
    toV = sample(1:nV,3) #selecting 3 nodes
    for (i in toV)
    {
       g1 = add.edges(g1,c(nV+1,i))
    }
    
  }else
  {
    indegree = degree(g1, mode = c("in"))
    pToBeCopied = indegree/nV
    pToBeCopied = head(pToBeCopied,-1)
    toV = sample(1:nV,size=3,prob=pToBeCopied)
    for (i in toV)
    {
       g1 = add.edges(g1,c(nV+1,i))
    }  
    
  }
  nV=nV+1
}
return(g1)
}
```


Let's generate one graph g1 
```{r}
g1 = generateGraph_singleOutDegree(0.5)

```


Let's show the log-log plot of the degree_distribution of g1 
and the cumulative degree distribution
```{r}
plot(degree_distribution(g1), log = "xy")
plot(degree_distribution(g1,cumulative = T), log = "xy")
```
As by the homework introduction and by the nature of the "preferential" attachement that implies power law.
also By looking at the plot we can tell that it is following a power law 
lets try fitting it into power law distribution using igraph method fit_power_law
```{r}
fit_power_law(degree_distribution(g1))
```
lets compare it to a simulated poisson distribution of similar paramters 
which reflects the differences between the two distributions
```{r}
g1_degree_dist = degree_distribution(g1)
plot(dpois(0:length(g1_degree_dist), lambda = 1), log="xy")
```

Now lets go deeper into exploring the factor of having "preferential" attachement 
Let generate graph g2 with gamma = 1 which will lead to a random network 

```{r}
g2 = generateGraph_singleOutDegree(1)
```

lets show the log-log plot of g2 degree distribution and cumulative degree distribution
the plots shows that it follows a poisson distribution

```{r}
plot(degree_distribution(g2))
plot(degree_distribution(g2), log = "xy")
plot(degree_distribution(g2,cumulative = T), log = "xy")
```


Lets try generating the graphs with out_degree = 3

```{r}
g3 = generateGraph_threeOutDegree(0.5)
```

Plotting the graphs 

```{r}
plot(degree_distribution(g3), log = "xy")
plot(degree_distribution(g3,cumulative = T), log = "xy")
```
It shows the similar distribution to g1, just "shifted" by 2 (the difference in the out_degrees)
And so the same rules with the poisson/power law distributions apply

lets draw g1 and g3 in_distributions to show the similarities 
```{r}
plot(degree_distribution(g1, cumulative = T, mode="in"),log = "xy")
plot(degree_distribution(g3, cumulative = T, mode="in"),log = "xy")
```

lets try fitting it into power law distribution using igraph method fit_power_law
```{r}
fit_power_law(degree_distribution(g3))
```

